Constraints on magic state protocols from the statistical mechanics of \\Wigner negativity

### 2021-06-28 magic_paper.tex

% Figures
% Refs
% Notation
% Proofreading --- Citations
% Proofreading --- Main text
% Proofreading --- Appendices

% Figs check --- consistency (size, font, text size, pads, axes, cbars), captions, folder

% Labels and Refs --- consistency, cref vs ref, thm vs lem vs prop, thm restatements, (??)

% Notation consistency --- noise vs error, \ln{} -> \ln, \ln vs \log vs \log_2, italics & \emph, Aberg -> {\AA}berg + [110 Knig], quasi-dist vs quasi-prob dist, def elbows, cumulant vs cumulative, stabilizer vs free, \p vs \w, W_\rho vs W_\rho(\bmz), \succ_r vs \succ
% UK vs US english

@@

\begin{abstract}
Magic states are key ingredients in schemes to realise universal fault-tolerant quantum computation.
Theories of magic states attempt to quantify this computational element via monotones and determine how these states may be efficiently transformed into useful forms. Here, we develop a statistical mechanical framework based on majorization in which Wigner negative magic states provide non-classical free energy that is processed under Clifford circuits for qudits of odd prime dimension. While traditional free energy is not defined in the presence of negativity, we show that majorization allows us to both quantify disorder in the Wigner representation and derive magic distillation bounds. These upper bounds are shown to be tighter than other bounds, such as from mana and thauma, and can incorporate hardware physics, such as temperature dependence and physical Hamiltonians. We also show that a subset of single-shot R\'{e}nyi entropies remain well-defined on quasi-distributions, are fully meaningful in terms of data processing and can acquire negative values that signal magic. We find that the mana of a magic state is the measure of divergence of these R\'{e}nyi entropies as one approaches the Shannon entropy for Wigner distributions, and discuss how distillation lower bounds could be obtained in this setting. This use of majorization for quasi-distributions could find application in other studies of non-classicality, and raises novel questions in the context of classical statistical mechanics.
\end{abstract}

@@

There has been rapid progress towards the goal of a fault-tolerant quantum computer~\cite{campbell_roads_2017, cit:raussendorf, Raussendorf_2013, Nickerson_2014, Nikahd_2017, chao_2018, lin_pieceable_2020, Lin_2020, Bourassa_2021}. However, many challenges remain and there is increasing need for theory to take into account physical limitations of the hardware involved. The surface code~\cite{Bravyi_1998, Freedman_2001, Dennis_2002, Raussendorf_2007} is a leading framework for fault-tolerance with very high error thresholds. Within this scheme, Clifford unitaries can be implemented in a robust, fault-tolerant way. However, due to the Eastin-Knill theorem~\cite{Eastin_2009}, we also know that it is impossible to have a universal set of transversal gates, and although Clifford unitaries can be realised transversally~\cite{Calderbank_1996, Steane_1996}, one needs to find ways around the Eastin-Knill restriction. This can be achieved by injecting in quantum states, called magic states, which promote the Clifford group to universal quantum computing~\cite{cit:bravyi}. The obstacle to this is that these states are invariably noisy, so protocols involving stabilizer operations must be employed to purify many copies of the magic states and improve the overall performance of the induced quantum gates~\cite{cit:bravyi, Jones_2013, Ogorman_2017}. A key question then arises about the overhead on purifying many copies of a magic state into less noisy forms.

@@

In this work, we develop a framework to analyse magic state distillation protocols where explicit physics of the system is incorporated in the distillation bounds. This is achieved by considering how a given magic state protocol transforms a \emph{pair} of quantum states -- one being a noisy magic state, and the other a stabilizer state that is distinguished by the physics of the system (for example a state at a characteristic temperature, or the maximally mixed state) and acts as a reference state for the protocol. The magic distillation bounds can then be expressed in terms of the physics of the reference state (e.g.~free energy changes in the case of temperature). Such bounds are of potential interest in assessing how physical features such as temperature, noise-biases or fixed-point structure associated with restricted gate-sets constrain distillation protocols~\cite{Tuckett_2019, Aliferis_2008, Stephens_2013, Li_2015, Babbush_2018, Guillaud_2019, Fowler_2019}.

The approach we take is most closely aligned with resource theories of magic, however it differs in key ways. We obtain distillation upper bounds without the use of monotones, but instead use tools from statistical mechanics and recent work in single-shot resource theories~\cite{cit:janzing, cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}. Our analysis relies on the discrete Wigner representation of quantum systems, in which all states and operations can be described on a discrete phase space~\cite{Ferrie_2008, Okay_2021}. Crucially, we focus on magic states with negativity in their Wigner representation, which is known to be a necessary condition for universality in the state-injection model~\cite{cit:veitch, cit:mari, cit:gottesman, cit:knill, Campbell_2011}. However, taking a statistical mechanical perspective in this context raises a problem: the standard Boltzmann entropy is not defined for quasi-distributions. We circumvent this obstacle by making use of a more fundamental tool -- majorization theory~\cite{cit:marshall, Veinott_1971, Ruch_1976}. To our knowledge majorization of quasi-distributions has not been considered in quantum physics before, and therefore these methods could find application in other studies of non-classicality in quantum systems~\cite{Fine_1982, Allahverdyan_2018, arvidsson_2020, halpern_2018, Lostaglio_2018, Levy_2020}.

\subsection{Overview of our results}
We begin by describing how stabilizer operations correspond to stochastic maps on a discrete phase space, and how majorization can be applied to constrain magic state transformations (Section~\ref{sec:ps}). We then consider families of magic protocols with increasing complexity. The simplest case we consider is for magic protocols that lead to unital channels (Section~\ref{sec:unital}). These are a subset of protocols that leave invariant some distinguished state or have some equilibrium fixed-point structure (Section~\ref{sec:frag}). We then consider protocols that are non-equilibrium processes, but only generate sub-linear correlations that enable a simple description in the thermodynamic limit (Section~\ref{sec:stab}), before giving entropic analysis that can be applied more generally (Section~\ref{section:entropies}).

@@

We provide explicit bounds for magic protocols that generate unital channels (Equation~(\ref{eq:unital-bound})), as well as bounds that incorporate the temperature and Hamiltonian of the system (Theorem~\ref{thm:free-energy}) and discuss extensions to more general scenarios. These bounds are shown to be tighter than other bounds such as those that come from mana and thauma (Section~\ref{sec:monotones_frag} and Figure~\ref{fig:distill_bounds}).

We find that the analysis in the presence of negativity displays a range of features that do not appear in classical statistical mechanics, and leads to a picture of Wigner negativity in a quantum circuit being described as non-classical free energy that is processed under stochastic dynamics. This is demonstrated by non-monotonic Lorenz curves (Section~\ref{sec:lc}), and by single-shot entropies for general Wigner quasi-distributions that are well-defined, obey data-processing under stabilizer operations and can take on negative values (Section~\ref{section:entropies}).

\begin{figure}[t]
    \centering
    \includegraphics[scale=0.35]{figs/distill_bounds.pdf}
    \caption{\textbf{Magic distillation bounds from majorization of Wigner quasi-distributions.} The above plot shows distillation upper bounds for qutrits obtained from majorisation on unital protocols. The constraints are plotted as a function of the depolarizing error $\epsilon$ for a noisy Strange state (Equation~\ref{eq:noisysn}). The bound $R_\infty$ follows from a Lorenz curve analysis presented in Section~\ref{sec:unital}, while the bound $R_{10}$ comes from single-shot entropies on the Wigner quasi-distributions, as discussed in Section~\ref{section:entropies}. Numerical bounds $R_{\rm{num}}$ are obtained by using the full set of majorization constraints. The new bounds $R_\infty$, $R_{10}$, $R_{\rm{num}}$ all provide stricter constraints than mana~\cite{cit:veitch2} and max--thauma $\theta_{\rm{max}}$~\cite{Wang_2020} bounds for noisy Strange states.
    }
    \label{fig:distill_bounds}
\end{figure}

We show that the mana of a quantum state arises as a residue in the $\alpha$--parameter for R\'{e}nyi entropies, and gives the rate of divergence to $-\infty$ as we approach the Shannon entropy of a Wigner distribution (Equation (\ref{eq:mana-divergence})).
Finally, in Section~\ref{sec:lower_bounds} we discuss how our approach could be extended to construct explicit lower bounds on distillation that take into account physical constraints.

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\cal{P}_d := \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\bmz \coloneqq (q, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\bmz} \coloneqq \tau^{q p} X^{q} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime~\cite{Appleby_2005}. The qubit case $d=2$ is known to have obstacles to using a Wigner representation to quantify magic in terms of negativity~\cite{Mermin_1990, cit:howard2}, although recent work~\cite{Raussendorf_qubits} has shown how to extend this to include qubits by enlarging the phase space. For a composite system with dimension $d_{\rm tot} = d_1 \dots d_n$ and Hilbert space $\cal{H}_{d_1} \otimes \dots \otimes \cal{H}_{d_n}$, we define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\bmz} \coloneqq D_{(q_1, p_1)} \otimes \dots \otimes D_{(q_n, p_n)},
\end{equation}
where now we have
\begin{align*}
	\bmz &\coloneqq (q_1, \dots, q_n, p_1, \dots, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n},
\end{align*}
to denote the phase space point for the composite system. 
For simplicity, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore we have that $\bmz \in \mathbb{Z}_d^{2n}$.

@@

The phase-point operators form an orthogonal operator basis with respect to the Hilbert-Schmidt inner product, as discussed in~\cref{app:wigner}.
Therefore, any quantum state $\rho$ can be expressed as a linear combination of them, $\rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz}$, where the coefficients $\W{\rho}(\z)$ give the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmz]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmz} \rho].
\end{equation}

For any quantum state $\rho$, the Wigner distribution $W_\rho(\bmz)$ is readily seen to be a $d^2$-dimensional quasi-probability distribution over $\cal{P}_d$. More precisely, $W_\rho(\bmz)$ is a real-valued function on $\P_d$ with the normalisation property $\sum_{\bmz} W_\rho(\bmz) = 1$ (see~\cref{app:wigner} for details).

@@

\subsection{Magic theories for quantum computation}
\label{sec:mono}
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/protocol_diagram.pdf}
    \caption{\textbf{Probing magic protocols with reference states.} 
	A magic protocol (top figure) converts $n$ copies of some noisy magic state $\rho$ with noise parameter $\epsilon$ to $m$ copies of some less noisy state $\rho'$ with $\epsilon' < \epsilon$. The physics of the protocol can be analysed by considering how it would transform some distinguished reference stabilizer state $\tau^{\otimes n}$ (bottom figure). For the case where this reference state is a thermal state at some temperature $T$ our magic distillation bounds are functions of $T$, the noise parameters $\epsilon, \epsilon'$, and free energies $F, F'$. In addition, a quantity $\phi$ (Equation~\ref{eq:phi}) relates the computational and energy bases, and corresponds to the degree to which the Wigner negativity of $\rho$ can be given a sharp energy via the system Hamiltonian.}
    \label{fig:sketch}
\end{figure}
A number of magic theories exist, where magic states are computational resource states with respect to a class of quantum operations that are considered free~\cite{Gour_2019}. One natural class of free operations are those obtained from Clifford unitaries, Pauli measurements and the ability to discard quantum systems. However, there are also other candidates~\cite{cit:ahmadi, cit:seddon, Wang_2019}.

@@

One prominent magic monotone is the \emph{mana} of a state~\cite{cit:veitch2}, defined as
\begin{equation}
    \M(\rho) \coloneqq \log (2\hspace{1pt}\sn{\rho}+1),
\end{equation}
where the \emph{sum-negativity}~\cite{cit:veitch2} is the sum of the negative components in $\W{\rho}$,
\begin{equation}
    \sn{\rho} \coloneqq \sum\limits_{\bmz: \W[\bmz]{\rho} < 0} \abs{\W[\bmz]{\rho}}.
\end{equation}
Using the fact that $W_\rho(\z)$ is a normalised quasi-distribution, we can also write $\M(\rho) = \sum_{\z} |W_\rho(\z)|$.

@@

In this work we develop bounds that apply to any reasonable magic theory. The central idea is to apply majorization theory to the quasi-probability distributions that arise for magic states. In the next section we explain how majorization relates to the free operations in a theory of magic.

@@

Any magic state protocol will correspond to a quantum channel $\E$, and so from Equation~(\ref{Wigner-Choi}) admits a Wigner representation $W_{\E}( \y | \bmz)$ that acts as a transition matrix mapping phase space points $\bmz \rightarrow \y$. The representation obeys the relation 
\begin{equation}
	W_{\E(\rho)} (\y) = \sum_{\bmz \in \P_d} W_{\E}( \y | \bmz) W_\rho(\bmz),
\end{equation}
for any $\rho$. Since the magic protocol sends free states to free states, $\E$ sends all positively represented quantum state to other positively represented quantum states. In such cases, it can be shown~\cite{Wang_2019} that if $\E$ is a free operation then the associated Wigner representation $W_{\E}(\y |\bmz)$ must form a stochastic matrix. In particular, all stabilizer operations correspond to stochastic matrices in the Wigner representation.

However, note that this mapping to stochastic maps is not one-to-one. Not all stochastic maps on the phase space correspond to valid quantum operations. The reason is that the maps must also respect the symplectic structure of the phase space, which is an additional non-trivial constraint.

@@

Majorisation~\cite{cit:marshall, Blackwell_1953} is a collection of powerful tools that has found many applications in quantum information theory~\cite{Nielsen_1999, cit:cwiklinski, cit:lostaglio2, cit:gour, cit:gour2, Horodecki_2003, Vallejos_2021}.
It describes the disorder of distributions that undergo stochastic transformations, and in its simplest form defines a pre-order on probability distributions. Given two distributions $\p= (p_1, \dots, p_n)$ and $\q = (q_1, \dots, q_n)$ over $n$ outcomes, we say that $\p$ majorises $\q$, denoted $\p \succ \q$, if there exists a bistochastic map $A = (A_{ij})$ such that $A\p = \q$, where bistochastic means that $A_{ij} \geq 0$ and $\sum_i A_{ij} = \sum_j A_{ij} = 1$. It can be shown~\cite{cit:marshall} that the condition $ \p \succ \q$ over probability distributions is equivalent to $n-1$ inequalities, which can be checked efficiently.

@@

In fact, we can further generalise to \emph{relative majorization}~\cite{Blackwell_1953, Ruch_1976, ruch_mixing_1978, Renes_2016, Buscemi_2017, Rethinasamy_2020}, defined as an ordering between pairs of vectors and write 
\begin{equation}
	(\p,\r) \succ (\q, \r')
\end{equation}
if and only if there is a stochastic map $A$ such that $A\r = \r'$ and $A\p = \q$. We retrieve $\bmd$--majorization when $\r = \r'$.

@@

Our analysis applies majorization to magic states at the level of the associated Wigner distributions. However, these are in general quasi-probability distributions, so it is important to check how majorization is computed for these cases and what differences quasi-distributions bring over genuine probability distributions.

@@

However, if $\w$ is a \emph{quasi-probability} distribution with negative values, and $\r$ a regular probability distribution things are different. Now the Lorenz curve is no longer monotonically increasing, but is a concave function that breaks through the $L_{\w|\r}(x) = 1$ barrier at an interior point and attains some non-trivial maximum $L_\star$ above the value $1$, before decreasing monotonically to $L_{\w|\r}(x)= 1$ at the end-point $x=1$. See~\cref{fig:lcs} for examples of non-monotonic Lorenz curves for quasi-distributions. Within quantum theory, the breaking of this barrier is associated with the degree of non-classicality.

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulant functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-probability distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho\coloneqq\rho_{\rm{S}}(\epsilon)^{\otimes n}$ (Equation~\ref{eq:noisysn}) for $n=2,4,6$ within $\R_{\sigma}$, where $\sigma = \id/3$. Solid lines represent pure Strange states, while dashed lines represent $\epsilon$-noisy Strange states with depolarising error $\epsilon = 0.1$.
    }
    \label{fig:lcs}
\end{figure}

Relative majorization is usually considered for probability distributions, so we need to verify that the same Lorenz curve conditions apply to quasi-probability distributions before proceeding with our analysis. This can be done from first principles, but a simpler way to see that this holds is to reduce to the problem of genuine distributions by first `masking' the negativity in the quasi-probability distribution $\w$ using the reference distribution $\r$ and then applying the conditions for relative majorization. This negativity masking then gives the following result.

>>>>

% Proofreading --- Citations (Aberg -> {\AA}berg)
% Proofreading --- Main text (Spacing)
% Proofreading --- Appendices

@@

\begin{abstract}
Magic states are key ingredients in schemes to realize universal fault-tolerant quantum computation.
Theories of magic states attempt to quantify this computational element via monotones and determine how these states may be efficiently transformed into useful forms. Here, we develop a statistical mechanical framework based on majorization in which Wigner negative magic states provide non-classical free energy that is processed under Clifford circuits for qudits of odd prime dimension. While traditional free energy is not defined in the presence of negativity, we show that majorization allows us to both quantify disorder in the Wigner representation and derive magic distillation bounds. These upper bounds are shown to be tighter than other bounds, such as from mana and thauma, and can incorporate hardware physics, such as temperature dependence and physical Hamiltonians. We also show that a subset of single-shot R\'{e}nyi entropies remain well-defined on quasi-distributions, are fully meaningful in terms of data processing and can acquire negative values that signal magic. We find that the mana of a magic state is the measure of divergence of these R\'{e}nyi entropies as one approaches the Shannon entropy for Wigner distributions, and discuss how distillation lower bounds could be obtained in this setting. This use of majorization for quasi-distributions could find application in other studies of non-classicality, and raises novel questions in the context of classical statistical mechanics.
\end{abstract}

@@

There has been rapid progress towards the goal of a fault-tolerant quantum computer~\cite{campbell_roads_2017, cit:raussendorf, Raussendorf_2013, Nickerson_2014, Nikahd_2017, chao_2018, lin_pieceable_2020, Lin_2020, Bourassa_2021}. Nevertheless, many challenges remain and there is increasing need for theory to take into account physical limitations of the hardware involved. The surface code~\cite{Bravyi_1998, Freedman_2001, Dennis_2002, Raussendorf_2007} is a leading framework for fault-tolerance with very high error thresholds. Within this scheme, Clifford unitaries can be implemented in a robust, fault-tolerant way. However, due to the Eastin-Knill theorem~\cite{Eastin_2009}, we also know that it is impossible to have a universal set of transversal gates, and although Clifford unitaries can be realized transversally~\cite{Calderbank_1996, Steane_1996}, one needs to find ways around the Eastin-Knill restriction. This can be achieved by injecting in quantum states, called magic states, which promote the Clifford group to universal quantum computing~\cite{cit:bravyi}. The obstacle to this is that these states are invariably noisy, so protocols involving stabilizer operations must be employed to purify many copies of the magic states and improve the overall performance of the induced quantum gates~\cite{cit:bravyi, Jones_2013, Ogorman_2017}. A key question then arises about the overhead on purifying many copies of a magic state into less noisy forms.

@@

In this work, we develop a framework to analyse magic state distillation protocols where explicit physics of the system is incorporated in the distillation bounds. This is achieved by considering how a given magic state protocol transforms a \emph{pair} of quantum states -- one being a noisy magic state, and the other a stabilizer state that is distinguished by the physics of the system (for example a state at a characteristic temperature, or the maximally mixed state) and acts as a reference state for the protocol. The magic distillation bounds can then be expressed in terms of the physics of the reference state (e.g.~free energy changes in the case of temperature). Such bounds are of potential interest in assessing how physical features such as temperature, noise biases or fixed-point structure associated with restricted gate-sets constrain distillation protocols~\cite{Tuckett_2019, Aliferis_2008, Stephens_2013, Li_2015, Babbush_2018, Guillaud_2019, Fowler_2019}.

The approach we take is most closely aligned with resource theories of magic, although it differs in key ways. We obtain distillation upper bounds without the use of monotones, but instead use tools from statistical mechanics and recent work in single-shot resource theories~\cite{cit:janzing, cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}. Our analysis relies on the discrete Wigner representation of quantum systems, in which all states and operations can be described on a discrete phase space~\cite{Ferrie_2008, Okay_2021}. Crucially, we focus on magic states with negativity in their Wigner representation, which is known to be a necessary condition for universality in the state-injection model~\cite{cit:veitch, cit:mari, cit:gottesman, cit:knill, Campbell_2011}. However, taking a statistical mechanical perspective in this context raises a problem: the standard Boltzmann entropy is not defined for quasi-distributions. We circumvent this obstacle by making use of a more fundamental tool -- majorization theory~\cite{cit:marshall, Veinott_1971, Ruch_1976}. To our knowledge majorization of quasi-distributions has not been considered in quantum physics before, and therefore these methods could find application in other studies of non-classicality in quantum systems~\cite{Fine_1982, Allahverdyan_2018, arvidsson_2020, halpern_2018, Lostaglio_2018, Levy_2020}.

\subsection{Overview of our results}
We begin by describing how stabilizer operations correspond to stochastic maps on a discrete phase space, and how majorization can be applied to constrain magic state transformations (Sec.~\ref{sec:ps}). We then consider families of magic protocols with increasing complexity. The simplest case we consider is for magic protocols that lead to unital channels (Sec.~\ref{sec:unital}). These are a subset of protocols that leave invariant some distinguished state or have some equilibrium fixed-point structure (Sec.~\ref{sec:frag}). We then consider protocols that are non-equilibrium processes, but only generate sub-linear correlations that enable a simple description in the thermodynamic limit (Sec.~\ref{sec:stab}), before giving entropic analysis that can be applied more generally (Sec.~\ref{section:entropies}).

@@

We provide explicit bounds for magic protocols that generate unital channels (Eq.~(\ref{eq:unital-bound})), as well as bounds that incorporate the temperature and Hamiltonian of the system (Theorem~\ref{thm:free-energy}) and discuss extensions to more general scenarios. These bounds are shown to be tighter than other bounds such as those that come from mana and thauma (Sec.~\ref{sec:monotones_frag} and Fig.~\ref{fig:distill_bounds}).

We find that the analysis in the presence of negativity displays a range of features that do not appear in classical statistical mechanics, and leads to a picture of Wigner negativity in a quantum circuit being described as non-classical free energy that is processed under stochastic dynamics. This is demonstrated by non-monotonic Lorenz curves (Sec.~\ref{sec:lc}), and by single-shot entropies for general Wigner quasi-distributions that are well-defined, obey data-processing under stabilizer operations and can take on negative values (Sec.~\ref{section:entropies}).

\begin{figure}[t]
    \centering
    \includegraphics[scale=0.35]{figs/distill_bounds.pdf}
    \caption{\textbf{Magic distillation bounds from majorization of Wigner quasi-distributions.} The above plot shows distillation upper bounds for qutrits obtained from majorization on unital protocols. The constraints are plotted as a function of the depolarizing error $\epsilon$ for a noisy Strange state (Eq.~(\ref{eq:noisysn})). The bound $R_\infty$ follows from a Lorenz curve analysis presented in Sec.~\ref{sec:unital}, while the bound $R_{10}$ comes from single-shot entropies on the Wigner quasi-distributions, as discussed in Sec.~\ref{section:entropies}. Numerical bounds $R_{\rm{num}}$ are obtained by using the full set of majorization constraints. The new bounds $R_\infty$, $R_{10}$, $R_{\rm{num}}$ all provide stricter constraints than mana~\cite{cit:veitch2} and max--thauma $\theta_{\rm{max}}$~\cite{Wang_2020} bounds for noisy Strange states.
    }
    \label{fig:distill_bounds}
\end{figure}

We show that the mana of a quantum state arises as a residue in the $\alpha$--parameter for R\'{e}nyi entropies, and gives the rate of divergence to $-\infty$ as we approach the Shannon entropy of a Wigner distribution (Eq.~(\ref{eq:mana-divergence})).
Finally, in Sec.~\ref{sec:lower_bounds} we discuss how our approach could be extended to construct explicit lower bounds on distillation that take into account physical constraints.

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\cal{P}_d \coloneqq \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\bmz \coloneqq (q, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\bmz} \coloneqq \tau^{q p} X^{q} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime~\cite{Appleby_2005}. The qubit case $d=2$ is known to have obstacles to using a Wigner representation to quantify magic in terms of negativity~\cite{Mermin_1990, cit:howard2}, although recent work~\cite{Raussendorf_qubits} has shown how to extend this to include qubits by enlarging the phase space. For a composite system with dimension $d_{\rm tot} = d_1 \dots d_n$ and Hilbert space $\cal{H}_{d_1} \otimes \dots \otimes \cal{H}_{d_n}$, we define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\bmz} \coloneqq D_{(q_1, p_1)} \otimes \dots \otimes D_{(q_n, p_n)},
\end{equation}
where now we have
\begin{align*}
	\bmz &\coloneqq (q_1, \dots, q_n, p_1, \dots, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n},
\end{align*}
to denote the phase space point for the composite system. 
For simplicity, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore we have that $\bmz \in \mathbb{Z}_d^{2n}$.

@@

The phase-point operators form an orthogonal operator basis with respect to the Hilbert-Schmidt inner product, as discussed in Appendix~\ref{app:wigner}.
Therefore, any quantum state $\rho$ can be expressed as a linear combination of them, $\rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz}$, where the coefficients $\W{\rho}(\z)$ give the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmz]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmz} \rho].
\end{equation}

For any quantum state $\rho$, the Wigner distribution $W_\rho(\bmz)$ is readily seen to be a $d^2$-dimensional quasi-distribution over $\cal{P}_d$. More precisely, $W_\rho(\bmz)$ is a real-valued function on $\P_d$ with the normalisation property $\sum_{\bmz} W_\rho(\bmz) = 1$ (see Appendix~\ref{app:wigner} for details).

@@

\subsection{Magic theories for quantum computation}
\label{sec:mono}
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/protocol_diagram.pdf}
    \caption{\textbf{Probing magic protocols with reference states.} 
	A magic protocol (top figure) converts $n$ copies of some noisy magic state $\rho$ with error parameter $\epsilon$ to $m$ copies of some less noisy state $\rho'$ with $\epsilon' < \epsilon$. The physics of the protocol can be analysed by considering how it would transform some distinguished reference stabilizer state $\tau^{\otimes n}$ (bottom figure). For the case where this reference state is a thermal state at some temperature $T$ our magic distillation bounds are functions of $T$, the error parameters $\epsilon, \epsilon'$, and free energies $F, F'$. In addition, a quantity $\phi$ (Eq.~(\ref{eq:phi})) relates the computational and energy bases, and corresponds to the degree to which the Wigner negativity of $\rho$ can be given a sharp energy via the system Hamiltonian.}
    \label{fig:sketch}
\end{figure}
A number of magic theories exist, where magic states are computational resource states with respect to a class of quantum operations that are considered free~\cite{Gour_2019}. One natural class of free operations are those obtained from Clifford unitaries, Pauli measurements and the ability to discard quantum systems. However, there are also other candidates~\cite{cit:ahmadi, cit:seddon, Wang_2019}.

@@

One prominent magic monotone is the \emph{mana} of a state~\cite{cit:veitch2}, defined as
\begin{equation}
    \M(\rho) \coloneqq \log (2\hspace{1pt}\sn{\rho}+1),
\end{equation}
where the \emph{sum-negativity}~\cite{cit:veitch2} is the sum of the negative components in $\W{\rho}$,
\begin{equation}
    \sn{\rho} \coloneqq \sum\limits_{\bmz: \W[\bmz]{\rho} < 0} \abs{\W[\bmz]{\rho}}.
\end{equation}
Using the fact that $W_\rho(\z)$ is a normalized quasi-distribution, we can also write $\M(\rho) = \sum_{\z} |W_\rho(\z)|$.

@@

In this work we develop bounds that apply to any reasonable magic theory. The central idea is to apply majorization theory to the quasi-distributions that arise for magic states. In the next section we explain how majorization relates to the free operations in a theory of magic.

@@

Any magic state protocol will correspond to a quantum channel $\E$, so from Eq.~(\ref{Wigner-Choi}) it admits a Wigner representation $W_{\E}(\y | \bmz)$ that acts as a transition matrix mapping phase space points $\bmz \rightarrow \y$. The representation obeys the relation 
\begin{equation}
	W_{\E(\rho)} (\y) = \sum_{\bmz \in \P_d} W_{\E}( \y | \bmz) W_\rho(\bmz),
\end{equation}
for any $\rho$. Since the magic protocol sends free states to free states, $\E$ sends all positively represented quantum state to other positively represented quantum states. In such cases, it can be shown~\cite{Wang_2019} that if $\E$ is a free operation then the associated Wigner representation $W_{\E}(\y |\bmz)$ must form a stochastic matrix. In particular, all stabilizer operations correspond to stochastic matrices in the Wigner representation.

We note that this mapping to stochastic maps is not one-to-one. Not all stochastic maps on the phase space correspond to valid quantum operations. The reason is that the maps must also respect the symplectic structure of the phase space, which is an additional non-trivial constraint.

@@

Majorisation~\cite{cit:marshall, Blackwell_1953} is a collection of powerful tools that has found many applications in quantum information theory~\cite{Nielsen_1999, cit:cwiklinski, cit:lostaglio2, cit:gour, cit:gour2, Horodecki_2003, Vallejos_2021}.
It describes the disorder of distributions that undergo stochastic transformations, and in its simplest form defines a pre-order on probability distributions. Given two distributions $\p= (p_1, \dots, p_n)$ and $\q = (q_1, \dots, q_n)$ over $n$ outcomes, we say that $\p$ majorizes $\q$, denoted $\p \succ \q$, if there exists a bistochastic map $A = (A_{ij})$ such that $A\p = \q$, where bistochastic means that $A_{ij} \geq 0$ and $\sum_i A_{ij} = \sum_j A_{ij} = 1$. It can be shown~\cite{cit:marshall} that the condition $ \p \succ \q$ over probability distributions is equivalent to $n-1$ inequalities, which can be checked efficiently.

@@

In fact, we can further generalize to \emph{relative majorization}~\cite{Blackwell_1953, Ruch_1976, ruch_mixing_1978, Renes_2016, Buscemi_2017, Rethinasamy_2020}, defined as an ordering between pairs of vectors and write 
\begin{equation}
	(\p,\r) \succ (\q, \r')
\end{equation}
if and only if there is a stochastic map $A$ such that $A\r = \r'$ and $A\p = \q$. We retrieve $\bmd$--majorization when $\r = \r'$.

@@

Our analysis applies majorization to magic states at the level of the associated Wigner distributions. Since these are in general quasi-distributions, it is important to check how majorization is computed for these cases and what differences quasi-distributions bring over genuine probability distributions.

@@

However, if $\w$ is a \emph{quasi-probability} distribution with negative values, and $\r$ a regular probability distribution things are different. Now the Lorenz curve is no longer monotonically increasing, but is a concave function that breaks through the $L_{\w|\r}(x) = 1$ barrier at an interior point and attains some non-trivial maximum $L_\star$ above the value $1$, before decreasing monotonically to $L_{\w|\r}(x)= 1$ at the end-point $x=1$. See Fig.~\ref{fig:lcs} for examples of non-monotonic Lorenz curves for quasi-distributions. Within quantum theory, the breaking of this barrier is associated with the degree of non-classicality.

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulative functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho\coloneqq\rho_{\rm{S}}(\epsilon)^{\otimes n}$ (Eq.~(\ref{eq:noisysn})) for $n=2,4,6$ within $\R_{\sigma}$, where $\sigma = \id/3$. Solid lines represent pure Strange states, while dashed lines represent $\epsilon$-noisy Strange states with depolarising error $\epsilon = 0.1$.
    }
    \label{fig:lcs}
\end{figure}

Relative majorization is usually considered for probability distributions, so we need to verify that the same Lorenz curve conditions apply to quasi-distributions before proceeding with our analysis. This can be done from first principles, but a simpler way to see that this holds is to reduce to the problem of genuine distributions by first `masking' the negativity in the quasi-distribution $\w$ using the reference distribution $\r$ and then applying the conditions for relative majorization. This negativity masking then gives the following result.
